---
title: "Osmia lignaria population genetic analysis"
author: "Jonathan Koch"
date: "2/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Volumes/GoogleDrive/My Drive/Manuscripts/Kochetal-OsmiaMicrosats/Data/PopulationGeneticAnalysis")
list.files()

library("poppr")
#monpop <- read.genalex("Osmia_PlexAll_Final_v4_sibsremoved.csv")
monpop <- read.genalex("Osmia_PlexAll_Final_v4_sibsremoved_14loci.csv")
# monpop <- read.genalex("1_Osmia_PlexAll_Final_v1.csv")

```

### The genotype accumulation curve ###

A genotype accumulation curve is a tool that allows you to assess how much power you have to discriminate between unique individuals given a random sample of n loci. We specified sample = 1000 in our function call. This means that for each boxplot, n loci were randomly sampled 1000 times in order to create the distribution. Since this data has been curated, we can see that we have reached the plateau with 4 loci. 

```{r echo=F, warning=F, message=F}

gac <- genotype_curve(monpop, sample = 1000, quiet = TRUE)

```
### Allele frequencies, missing data, and ploidy ###

A good first step after loading data is to look for missing data, rare alleles and overall quality of your data:

```{r echo=F, warning=F, message=F}

(pinflt <- locus_table(monpop))
CR.monpop <- as.data.frame((pinflt <- locus_table(monpop, population = c("Cub River"))))
BC.monpop <- as.data.frame((pinflt <- locus_table(monpop, population = c("Birch Creek"))))
missing <- info_table(monpop, type = "missing")
#write.csv(pinflt, "Osmia lignaria locus summary.csv")
#write.csv(missing, "Osmia lignaria missing data summary.csv")

# Compare 1-D
var.test(CR.monpop$`1-D`, BC.monpop$`1-D`) # F = 1.462, num df = 14, denom df = 14, p-value = 0.4864
t.test(CR.monpop$`1-D`, BC.monpop$`1-D`, var.equal = TRUE) # t = -0.35309, df = 28, p-value = 0.7267
mean(BC.monpop$`1-D`)
sd(BC.monpop$`1-D`)/sqrt(nrow(BC.monpop))
mean(CR.monpop$`1-D`)
sd(CR.monpop$`1-D`)/sqrt(nrow(CR.monpop))
```

### Removing loci and genotypes ##

When removing loci or genotypes, you can specify a cutoff representing the percent missing to be removed. The default is 0.05 (5%).

```{r echo=F, warning=F, message=F}
library("magrittr")
monpop %>% missingno("geno", cutoff = 0.20) %>% info_table(plot = TRUE)
# monpop %>% missingno("loci", cutoff = 0.50) %>% info_table(plot = TRUE, scale = TRUE)

```
```{r}
# strata G
# make sure you have Rtools installed
# install from GitHub
# devtools::install_github('ericarcher/strataG', build_vignettes = TRUE)

library(strataG)
library(adegenet)

# read loci
df <- read.csv("Osmia_PlexAll_Final_v4_adegenet_sibsremoved.csv", header = TRUE)
df.2 <- df2gtypes(df, ploidy = 2, id.col = 1, strata.col = 2, loc.col = 3)

# test for Hardy-Weinberg Equilibrium
hwe_df.2 <- hweTest(df.2)
hwe_df.2 <- as.data.frame(hwe_df.2)
hwe_df.2$p.value.adjust <- p.adjust(hwe_df.2$hwe_df.2, method = "bonferroni")
# write.csv(hwe_df.2, "Results/results_hwe_Osmialignaria_BCCR.csv", row.names = TRUE)

# test for Linkage Disequilibirum
LD_df.2 <- LDgenepop(df.2)
LD_df.2
LD_df.2.adjust <- p.adjust(LD_df.2$p.value)
LD_df.2$p.value.adjust <- LD_df.2.adjust
(sort(LD_df.2$p.value.adjust))
# write.csv(LD_df.2, "Results/results_LD_Osmialignaria_BCCR.csv", row.names = TRUE)

# Remove loci not in HWE or in LD
ncol(df)
getLociNames(df.2)
df.3 <- subset(df, select = -c(Oli053, Oli064, Oli076,Oli101,
                               Oli127, Oli160,Oli61901,Oli84207,
                               Oli053.1, Oli064.1, Oli076.1,Oli101.1,
                               Oli127.1, Oli160.1,Oli61901.1,Oli84207.1))
ncol(df.3)

df.4 <- df2gtypes(df.3, ploidy = 2, id.col = 1, strata.col = 2, loc.col = 3)

```

```{r}
# genetic diversity
# expected heterozygosity
he <- strataG::heterozygosity(df.4, by.strata = TRUE,
                                type = c("expected"))
ho <- strataG::heterozygosity(df.4, by.strata = TRUE,
                                type = c("observed"))
# Merge the data
heho <- merge(he,ho)

# Birch Creek He vs Ho
BC.df <- heho %>%
  subset(stratum=="Birch Creek")
var.test(BC.df$exptd.het, BC.df$obsvd.het) # F = 0.61897, num df = 13, denom df = 13, p-value = 0.3984
t.test(BC.df$exptd.het, BC.df$obsvd.het, var.equal = TRUE) # t = 0.32483, df = 24.635, p-value = 0.7481

# Cub River He vs Ho
CR.df <- heho %>%
  subset(stratum=="Cub River")
var.test(CR.df$exptd.het, CR.df$obsvd.het) # F = 0.97675, num df = 13, denom df = 13, p-value = 0.9668
t.test(CR.df$exptd.het, CR.df$obsvd.het, var.equal = TRUE) # t = 0.70948, df = 25.996, p-value = 0.4843

# Birch Creek vs Cub River He
var.test(he$exptd.het~he$stratum) # F = 0.69418, num df = 13, denom df = 13, p-value = 0.5197
t.test(he$exptd.het~he$stratum, var.equal = TRUE) # t = 0.37863, df = 26, p-value = 0.708

# allelic richness
all.rich <- strataG::allelicRichness(df.4, by.strata = TRUE)
var.test(all.rich$allelic.richness~all.rich$stratum) # F = 0.68377, num df = 13, denom df = 13, p-value = 0.5027
t.test(all.rich$allelic.richness~all.rich$stratum, var.equal = TRUE) # t = 0.79038, df = 26, p-value = 0.4365



# Number of alleles
num.alleles <- numAlleles(df.4, by.strata = TRUE)
var.test(num.alleles$num.alleles~num.alleles$stratum) # F = 0.38337, num df = 13, denom df = 13, p-value = 0.09586
t.test(num.alleles$num.alleles~num.alleles$stratum, var.equal = TRUE) # t = -0.26198, df = 26, p-value = 0.7954

# proportion of unique alleles
uniq.alleles <- propUniqueAlleles(df.4, by.strata = TRUE)
var.test(uniq.alleles$num.unique~uniq.alleles$stratum) # F = 0.31422, num df = 13, denom df = 13, p-value = 0.04603
t.test(uniq.alleles$num.unique~uniq.alleles$stratum, var.equal = FALSE) # t = -0.20208, df = 20.436, p-value = 0.8418
wilcox.test(uniq.alleles$num.unique~uniq.alleles$stratum) # W = 112.5, p-value = 0.5129

num.unique.BC <- uniq.alleles %>%
  subset(stratum == "Birch Creek")
median(num.unique.BC$num.unique)

num.unique.CR <- uniq.alleles %>%
  subset(stratum == "Cub River")
median(num.unique.CR$num.unique)

wilcox.test(uniq.alleles$num.unique~uniq.alleles$stratum) #  W = 112.5, p-value = 0.5129

# one big table of values
gd.table <- cbind(he,ho$obsvd.het, all.rich$allelic.richness, num.alleles$num.alleles, uniq.alleles$num.unique, test$`1-D`)

# write.csv(gd.table, "Results/Osmialignaria_genetic_diversity_metrics.csv")
```

```{r}
library(inbreedR)

# convert data

df.3.BC <- subset(df.3, df.3$Pop=="Birch Creek")
df.3.CR <- subset(df.3, df.3$Pop=="Cub River")

df.3.BC <- subset(df.3.BC, select = -c(me, Pop))
df.3.CR <- subset(df.3.CR, select = -c(me, Pop))

df.4.BC <- convert_raw(df.3.BC)
df.4.CR <- convert_raw(df.3.CR)

check_data(df.4.BC, num_ind = 18, num_loci = 14)
check_data(df.4.CR, num_ind = 23, num_loci = 14)


g2_all_microsats.BC <- g2_microsats(df.4.BC, nperm = 100, nboot = 100, CI=0.95) 
g2_all_microsats.BC # No evidence of inbreeding
g2_all_microsats.CR <- g2_microsats(df.4.CR, nperm = 100, nboot = 100, CI=0.95)
g2_all_microsats.CR # Evidence of inbreeding

par(mfrow=c(2,1))

plot(g2_all_microsats.BC, main = "Birch Creek",
     col = "cornflowerblue", cex.axis=0.85)
plot(g2_all_microsats.CR, main = "Cub River",
     col = "cornflowerblue", cex.axis=0.85)

HHC.BC <- HHC(df.4.BC, reps = 1000)
HHC.BC
HHC.CR <- HHC(df.4.CR, reps = 1000)
HHC.CR

par(mfrow=c(1,2))
plot(HHC.BC, main = "Birch Creek",
     col = "cornflowerblue", cex.axis=0.85)
plot(HHC.CR, main = "Cub River",
     col = "darkgoldenrod1", cex.axis=0.85)
```

```{r}
ovl <- overallTest(df.4, stats = c("all"), nrep = 1000, keep.null = TRUE) # significant Fst

popStruct <- popStructTest(df.4, stats = c("all"), nrep = 1000, quietly = TRUE)
popStruct

```

```{r}
# DAPC requires the adegenet package. Let's load this package:
library("adegenet")
pop(monpop) <- monpop$pop
dapc.monpop <- dapc(monpop, var.contrib = TRUE, scale = FALSE, n.pca = 10, n.da = nPop(monpop)-1)
scatter(dapc.monpop, cell = 0, pch = 18:19, cstar = 0, mstree = TRUE, lwd = 2, lty = 2)
```

### AMOVA
```{r}

table(strata(monpop))

Aeutamova <- poppr.amova(monpop, ~Pop, cutoff = 0.20) # panmictic population; most variation occurs within samples.
set.seed(1999)
Aeutamova
Aeutsignif<- randtest(Aeutamova, nrepet = 999)
Aeutsignif
plot(Aeutsignif)
```

```{r}
library("ape") # To visualize the tree using the "nj" function
library("magrittr")

micdist <- provesti.dist(monpop)
theTree <- micdist %>%
  nj() %>%    # calculate neighbor-joining tree
  ladderize() # organize branches by clade
plot(theTree)
add.scale.bar(length = 0.05) # add a scale bar showing 5% difference.

```

